---
title: "Plots by individual country"
output:
  html_document:
    df_print: paged
    toc: yes
---


## Packages and functions

Load required libraries:

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(viridis)
library(ggplot2)
```

Other custom functions:
```{r, echo=F, message=F, warning=F}
### not in
'%!in%' <- function(x,y)!('%in%'(x,y))

```

Custom colors:
```{r custom colors, echo=F, message=F, warning=F}
## IUCN official colors
# Assuming order of levels is: "re", "cr", "en", "vu", "nt", "lc", "dd", "not_assessed", "unknown" (for regional, and w/o "re" for global). Make sure to change the levels to that order before plotting.
IUCNcolors<-c("brown2", "darkorange", "yellow", "green", "darkgreen", "darkgrey", "azure2", "bisque1")
IUCNcolors_regional<-c("darkorchid2", "brown2", "darkorange", "yellow", "green", "darkgreen", "darkgrey", "azure2", "bisque1")

## nice soft ramp for taxonomic groups
taxoncolors<-cividis(12) # same than using cividis(length(levels(as.factor(metadata$taxonomic_group))))
  
```

## Get data 

Get clean data (output of step 3) ready to estimate the indicators:

```{r, echo=TRUE}
# Get data:
indicators_full<-read.csv(file="indicators_full.csv")
indNe_data<-read.csv("indNe_data.csv")
```


```{r, echo=F, message=F, warning=F}
# Optional step if you have several countries in the dataset

# Select which country you want to analyse. For this after "desired_country" in the code below, write the name of your country exactly as it is written in kobo (English lowercase, no spaces use "_" instead)

desired_country<-"mexico"

# Subset data to desired country:
indNe_data<-indNe_data %>% filter(country_assessment==desired_country)
indicators_full<-indicators_full %>% filter(country_assessment==desired_country)

```

Which country is being analysed?
```{r, echo=F, message=F, warning=F}
desired_country<-unique(indicators_full$country_assessment)
desired_country
```


## General description of the dataset

`r desired_country` has a total of `r nrow(indicators_full)` records. Of those, some taxa could have been assessed more than once, for example to account for different methods to define populations.


```{r, echo=F, message=F, warning=F}
# To explore what kind of taxa the country assessed regardless of if they assessed them once or more, lets create a dataset keeping all single assessed taxa, plus only the first assessment for taxa assessed multiple times. 

# object with single assessed taxa, plus only the first assessment for taxa assessed multiple times
firstmulti<-indicators_full[!duplicated(cbind(indicators_full$taxon, indicators_full$country_assessment)), ]

```

How many taxa were assessed (i.e. counting only once taxa that were assessed multiple times)?

```{r, echo=F, message=F, warning=F}
# how many?
nrow(firstmulti)
```

Which taxa where assessed more than once? (if any)

```{r, echo=F, message=F, warning=F}
indicators_full[duplicated(cbind(indicators_full$taxon, indicators_full$country_assessment)), ]
```


### Taxonomic groups and endemicity


```{r, echo=F, message=F, warning=F, out.width="700px", out.height="400px"}
ggplot(indicators_full, aes(x=taxonomic_group, fill=national_endemic)) + 
  geom_bar(stat = "count") +
  xlab("") +
  ggtitle("Number of taxa by taxonomic groups and endemicity") +
  theme_light() +
  theme(legend.position="bottom")

```
### Proportion of species distribution within the country

```{r, echo=F, message=F, warning=F, out.width="700px", out.height="400px"}
ggplot(indicators_full, aes(x=taxonomic_group, fill=country_proportion)) + 
  geom_bar(stat = "count") +
  xlab("") +
  ggtitle("Number of taxa by taxonomic groups and proportion \nof taxon distribution within the country") +
  theme_light() +
  theme(legend.position="bottom")

```

### Realm and habitat

Plot by Realm

```{r, echo=F, message=F, warning=F}
ggplot(indicators_full, aes(x=taxonomic_group, fill=realm)) + 
  geom_bar(stat = "count") +
  xlab("") +
  ggtitle("Number of taxa by taxonomic groups and realm") +
  theme_light() +
  theme(legend.position="bottom")
```

Habytat type as IUCN categories (a species can inhabit more than one type of habitat, that is why there are so many groups):


```{r, echo=F, message=F, warning=F}
ggplot(indicators_full, aes(x=taxonomic_group, fill=IUCN_habitat)) + 
  geom_bar(stat = "count") +
  xlab("") +
  ggtitle("Number of taxa by taxonomic groups and habitat") +
  theme_light() +
  theme(legend.position="bottom")
```


### Population size data availability

Population size (Nc or Ne) data availability at the taxon level:

```{r, echo=F, message=F, warning=F}
ggplot(indicators_full, aes(x=taxonomic_group, fill=popsize_data)) + 
  geom_bar(stat = "count") +
  coord_flip() +
  scale_fill_manual(values=c("#2ca02c", "#1f77b4", "grey80"),
                    breaks=c("yes", "data_for_species", "insuff_data_species"),
                    labels=c("Population level", "Species or subspecies level", "Insufficient data")) +
  labs(fill="Population size data availability",
       x="",
       y="Number of taxa (including records of taxa assessed more than once)") +
  theme_light() +
  theme(panel.border = element_blank(), legend.position="top")
  
```

Ne available?

```{r, echo=F, message=F, warning=F}
indicators_full %>% 
  filter(!is.na(ne_pops_exists)) %>% 
  filter(ne_pops_exists!="other_genetic_info") %>%
    ggplot(aes(x=taxonomic_group, fill=ne_pops_exists)) + 
  geom_bar() +
  coord_flip() +
scale_fill_manual(labels=c("no", "yes"),
                      breaks=c("no_genetic_data", "ne_available"),
                      values=c("#ff7f0e", "#2ca02c")) +
xlab("") +
ylab("Number of taxa") +
labs(fill="Ne available \n(from genetic data)")  +
theme_light() +
theme(text = element_text(size = 14), legend.position = "right", panel.border = element_blank())


```

Nc data available by taxa?

```{r, echo=F, message=F, warning=F}
indicators_full %>%
  filter(!is.na(nc_pops_exists)) %>%
    ggplot(aes(x=taxonomic_group, fill=nc_pops_exists)) +
    geom_bar() +
    coord_flip() +
    scale_fill_manual(values=c("#ff7f0e", "#2ca02c")) + 
    labs(fill="Nc available") +
    xlab("") +
    ylab("Number of taxa") +
    theme_light() +
    theme(text = element_text(size = 14), legend.position = "right", panel.border = element_blank())
```

What kind of Nc data?

```{r, echo=F, message=F, warning=F}
indNe_data %>%
  filter(!is.na(NcType)) %>%
  ggplot(aes(x=taxonomic_group, fill=NcType))+
  geom_bar() +
  scale_fill_manual(labels=c("Point", "Range \nor qualitative", "Unknown"),
                      breaks=c("Nc_point", "Nc_range", "unknown"),
                      values=c("#0072B2", "#E69F00", "grey80")) +
  xlab("") +
  ylab("Number of populations") +
  coord_flip() +
  labs(fill="Type of Nc data \nby population") +
  theme_light() +
  theme(text = element_text(size = 14), legend.position = "right", panel.border = element_blank())
 
```



## Methods used to define populations
Frequency table of methods used to define population
```{r, echo=F, message=F, warning=F}
table(indicators_full$defined_populations)
```

### Plot number of populations by method
```{r, echo=F, message=F, warning=F, out.height="500px", out.width="1064px"}
# Prepare data for plot with nice labels:
# sample size of TOTAL populations
sample_size <- indicators_full %>%
                    filter(!is.na(indicatorPM)) %>% 
                    group_by(defined_populations) %>% summarize(num=n())

# custom axis
## new dataframe
df<-indicators_full %>% 
  filter(!is.na(indicatorPM)) %>%
  filter(n_extant_populations<500) %>%
    # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(defined_populations, " (n= ", num, ")"))


# plot for number of pops
df %>%
  ggplot(aes(x=myaxis, y=n_extant_populations, color=defined_populations)) +
          geom_boxplot() + xlab("") + ylab("Number of mantained populations") +
          geom_jitter(size=.4, width = 0.1, color="black") +
  coord_flip() +
  theme_light() +
  theme(panel.border = element_blank(), legend.position="none",
        plot.margin = unit(c(0, 0, 0, 0), "cm")) + # this is used to decrease the space between plots
  scale_x_discrete(limits=rev) +  
  theme(text = element_text(size = 13))

```


## Missing data on extant and extinct populations

We have NA in indicator PM because in some cases the number of extinct populations is unknown, therefore the operation cannot be computed. 

### Counts
Total records with NA in extant populations:
```{r, echo=F, message=F, warning=F}
sum(is.na(indicators_full$n_extant_populations))
```

Taxa with NA in extant populations:
```{r, echo=F, message=F, warning=F}
indicators_full %>%
  filter(is.na(n_extant_populations)) %>%
    select(taxonomic_group, taxon, n_extant_populations, n_extinct_populations)
```

Total taxa with NA in **extinct** populations:
```{r, echo=F, message=F, warning=F}
sum(is.na(indicators_full$n_extinct_populations))
```

Do taxa with NA for extant also have NA for extinct?

```{r, echo=F, message=F, warning=F}
indicators_full$taxon[is.na(indicators_full$n_extant_populations)] %in% indicators_full$taxon[is.na(indicators_full$n_extinct_populations)]
```

### Plot missing data extinct populations
```{r, echo=F, message=F, warning=F}
indicators_full %>%
  ggplot(aes(x=taxonomic_group, fill=is.na(n_extinct_populations))) +
  geom_bar() +
  scale_fill_manual(labels=c("number of populations known", "missing data"),
                    values=c("#2ca02c", "#ff7f0e")) + 
  coord_flip() +
  labs(fill="Extinct populations") +
  xlab("") + ylab("Number of taxa") +
  theme_light() +
  theme(panel.border = element_blank())
```


## Ne > 500 indicator

In total, there was data to compute the Ne indicator in the following number of taxa:

```{r}
sum(!is.na(indicators_full$indicatorNe))
```

Considering those `r sum(!is.na(indicators_full$indicatorNe))` taxa, the Ne indicator mean value for the country is:

```{r, echo=F, message=F, warning=F}
mean_ne <- mean(indicators_full$indicatorNe, na.rm = TRUE)
mean_ne
```

### Ne indicator by taxon 

```{r, echo=F, message=F, warning=F}
indicators_full %>%
  filter(!is.na(indicatorNe)) %>%
  
ggplot(aes(x = indicatorNe, y = reorder(taxon, indicatorNe))) +
  geom_col(fill = "#0B2B4B") +
  geom_vline(xintercept = mean_ne,
             colour = "orange",
             linetype = "dashed",
             linewidth = 1) +
  geom_text(aes(label = round(indicatorNe, 2),
                x = indicatorNe + 0.02),                 # nudge labels slightly right
            size = 3,
            hjust = 0) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02)),
                     limits = c(0, 1)) +                 # adjust if indicatorNe is 0–1
  labs(
    x = "Proportion of populations within species > 500 Ne",
    y = "",
    title = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10)
  )
```


### Ne > 500  by type of range

By type of range in the entire dataset:

```{r, echo=F, message=F, warning=F}
# get sample size by desired category
sample_size <- indicators_full  %>%
                    filter(!is.na(indicatorNe)) %>% 
                    group_by(species_range) %>% summarize(num=n())

# plot
indicators_full  %>% 
  # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(species_range, " (n= ", num, ")")) %>%

  # plot
  ggplot(aes(x=myaxis, y=indicatorNe , fill=species_range)) +
      geom_violin(width=1, linewidth = 0)  +
      geom_jitter(size=.5, width = 0.1) +
      xlab("") + ylab("Proportion of popuations with Ne > 500") +
      coord_flip() +
      theme_light() +
      theme(panel.border = element_blank(), legend.position="none", text= element_text(size=20))
```

### Ne > 500 by IUCN status

By global IUCN:

```{r indicatorNe gobalIUCN, echo=F, message=F, warning=F}

## Global IUCN
## prepare data
# add sampling size
sample_size <- indicators_full %>%
               filter(!is.na(indicatorNe)) %>% 
               group_by(global_IUCN) %>% summarize(num=n())

# new df 
df<- indicators_full %>% 
     filter(!is.na(indicatorNe)) %>% 
        # add sampling size 
        left_join(sample_size) %>%
        mutate(myaxis = paste0(global_IUCN, " (n= ", num, ")"))


# change order of levels so that they are in the desired order
df$myaxis<-factor(df$myaxis, 
                  #grep is used below to get the sample size, which may change depending on the data
                  levels=c(grep("cr", unique(df$myaxis), value = TRUE),
                          grep("en", unique(df$myaxis), value = TRUE),
                          grep("vu", unique(df$myaxis), value = TRUE),
                          grep("nt", unique(df$myaxis), value = TRUE),
                          grep("lc", unique(df$myaxis), value = TRUE),
                          grep("dd", unique(df$myaxis), value = TRUE),
                          grep("not_assessed", unique(df$myaxis), value = TRUE),
                          grep("unknown", unique(df$myaxis), value = TRUE)))

df$global_IUCN<-factor(df$global_IUCN, levels=c("cr", "en", "vu", "nt", "lc", "dd", "not_assessed", "unknown"))


      
# plot
df %>%
    ggplot(aes(x=myaxis, y=indicatorNe , fill=global_IUCN)) +
      geom_violin(width=1.5, linewidth = 0)  +
      geom_jitter(size=.5, width = 0.1) +
      xlab("") + ylab("Proportion of popuations with Ne > 500") +
      coord_flip() +
      scale_fill_manual(values= IUCNcolors, # iucn color codes
                        breaks=c(levels(df$global_IUCN))) +
      theme_light() +
      ggtitle("global Red List") +
      theme(panel.border = element_blank(), legend.position="none", text= element_text(size=15))

```

By regional IUCN (this may make no sense for your country if there is no IUCN redlist assessments at regional level:

```{r, echo=F, message=F, warning=F}
## Regional IUCN

## prepare data
# add sampling size
sample_size <- indicators_full %>%
               filter(!is.na(indicatorNe)) %>% 
               group_by(regional_redlist) %>% summarize(num=n())

# new df 
df<- indicators_full %>% 
     filter(!is.na(indicatorNe)) %>% 
        # add sampling size 
        left_join(sample_size) %>%
        mutate(myaxis = paste0(regional_redlist, " (n= ", num, ")"))

# change order of levels so that they are in the desired order
df$myaxis<-factor(df$myaxis, 
                  #grep is used below to get the sample size, which may change depending on the data
                  levels=c(grep("re", unique(df$myaxis), value = TRUE),
                          grep("cr", unique(df$myaxis), value = TRUE),
                          grep("en", unique(df$myaxis), value = TRUE),
                          grep("vu", unique(df$myaxis), value = TRUE),
                          grep("nt", unique(df$myaxis), value = TRUE),
                          grep("lc", unique(df$myaxis), value = TRUE),
                          grep("dd", unique(df$myaxis), value = TRUE),
                          grep("not_assessed", unique(df$myaxis), value = TRUE),
                          grep("unknown", unique(df$myaxis), value = TRUE)))

df$regional_redlist<-factor(df$regional_redlist, levels=c("re","cr", "en", "vu", "nt", "lc", "dd", "not_assessed", "unknown"))
      
# plot
df %>%
    ggplot(aes(x=myaxis, y=indicatorNe , fill=regional_redlist)) +
      geom_violin(width=1, linewidth = 0)  +
      geom_jitter(size=.5, width = 0.1) +
      xlab("") + ylab("Proportion of popuations with Ne > 500") +
      coord_flip() +
      scale_fill_manual(values= IUCNcolors_regional, # iucn color codes
                        breaks=c(levels(df$regional_redlist))) +
      theme_light() +
      ggtitle("regional Red List") +
      theme(panel.border = element_blank(), legend.position="none", text= element_text(size=15))



```
### Ne > 500 by endemicity
```{r, echo=F, message=F, warning=F}
# get sample size by desired category
sample_size <- indicators_full  %>%
                    filter(!is.na(indicatorNe)) %>% 
                    group_by(national_endemic) %>% summarize(num=n())

# plot
indicators_full  %>% 
  # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(national_endemic, " (n= ", num, ")")) %>%

  # plot
  ggplot(aes(x=myaxis, y=indicatorNe , fill=national_endemic)) +
      geom_violin(width=1, linewidth = 0)  +
      geom_jitter(size=.5, width = 0.1) +
      xlab("") + ylab("Proportion of popuations with Ne > 500") +
      coord_flip() +
      theme_light() +
      theme(panel.border = element_blank(), legend.position="right", text= element_text(size=20))
```


### Distribution of Ne values

How is Ne data distributed? 

```{r, echo=F, message=F, warning=F}
summary(indNe_data$Ne)
```


Boxplot of Ne values:
```{r, echo=F, message=F, warning=F}
indNe_data %>%
  ggplot(aes(x=taxonomic_group, y=Ne)) +
  geom_boxplot() + geom_point(aes(x=taxonomic_group, y=Ne))+
theme_light()
```

Check outliers (Ne very high):

```{r, echo=F, message=F, warning=F}
indNe_data %>% 
  filter(Ne > 100000) %>%
  select(country_assessment, name_assessor, taxon, taxonomic_group, Ne, NeLower, NeUpper, multiassessment, population)
```

Ne Boxplot filtering outliers (if any)
```{r, echo=F, message=F, warning=F}
indNe_data %>% filter(Ne < 100000) %>%
  ggplot(aes(x=taxonomic_group, y=Ne)) +
  geom_boxplot() + geom_point(aes(x=taxonomic_group, y=Ne))+
theme_light()
```

### Distribution of Nc values and ranges

Point data
```{r, echo=F, message=F, warning=F}
indNe_data %>%
  ggplot(aes(x=taxonomic_group, y=NcPoint)) +
  geom_boxplot() + geom_point(aes(x=taxonomic_group, y=NcPoint))+
theme_light()

```

Range data (categorical):

```{r, echo=F, message=F, warning=F}
# re-order levels
backupt<-indNe_data
indNe_data$NcRange<-as.factor(indNe_data$NcRange)
table(indNe_data$NcRange)
indNe_data$NcRange<-factor(indNe_data$NcRange, levels = c("less5000_bymuch", "less_5000", "range_includes_5000",
                                "more_5000", "more_5000_bymuch"))

indNe_data %>%
  filter(!is.na(NcRange)) %>%
  ggplot(aes(x=taxonomic_group, fill=NcRange)) +
  geom_bar(position = "stack") +
  coord_flip() +
  scale_fill_manual(breaks = c("less5000_bymuch", "less_5000", "range_includes_5000",
                                "more_5000", "more_5000_bymuch"),
                    values = c("darkred", "red",
                                 "orange",
                                 "green", "darkgreen")) +
theme_light()
```




## Proportion of populations mantained (PM indicator)

### Plot Proportion of populations mantained (PM indicator) by method
```{r, echo=F, message=F, warning=F}
# Prepare data for plot with nice labels:
# sample size of TOTAL populations
sample_size <- indicators_full %>%
                    filter(!is.na(indicatorPM)) %>% 
                    group_by(defined_populations) %>% summarize(num=n())

# custom axis
## new dataframe
df<-indicators_full %>% 
  filter(n_extant_populations<500) %>%
  filter(!is.na(indicatorPM)) %>%
    # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(defined_populations, " (n= ", num, ")"))



## plot for indicator 2
df %>%
  filter(n_extant_populations<500) %>%
  ggplot(aes(x=myaxis, y=indicatorPM, color=defined_populations)) +
          geom_boxplot() + xlab("") + ylab("Proportion of maintained populations") +
          geom_jitter(size=.4, width = 0.1, color="black") +
  coord_flip() +
  theme_light() +
  theme(panel.border = element_blank(), legend.position="none") +
  scale_x_discrete(limits=rev) +
  theme(text = element_text(size = 13))


```

### by type of range

By type of range in the entire dataset:
  
```{r, echo=F, message=F, warning=F}
# get sample size by desired category
sample_size <- indicators_full  %>%
  filter(!is.na(indicatorPM)) %>% 
  group_by(species_range) %>% summarize(num=n())

# plot
indicators_full  %>% 
  # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(species_range, " (n= ", num, ")")) %>%
  
  # plot
  ggplot(aes(x=myaxis, y=indicatorPM , fill=species_range)) +
  geom_violin(width=1, linewidth = 0)  +
  geom_jitter(size=.5, width = 0.1) +
  xlab("") + ylab("Proportion of pmaintained populations") +
  coord_flip() +
  theme_light() +
  theme(panel.border = element_blank(), legend.position="none", text= element_text(size=20))
```

###  by IUCN status

By global IUCN:

```{r, echo=F, message=F, warning=F}

## Global IUCN
## prepare data
# add sampling size
sample_size <- indicators_full %>%
  filter(!is.na(indicatorPM)) %>% 
  group_by(global_IUCN) %>% summarize(num=n())

# new df 
df<- indicators_full %>% 
  filter(!is.na(indicatorPM)) %>% 
  # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(global_IUCN, " (n= ", num, ")"))


# change order of levels so that they are in the desired order
df$myaxis<-factor(df$myaxis, 
                  #grep is used below to get the sample size, which may change depending on the data
                  levels=c(grep("cr", unique(df$myaxis), value = TRUE),
                           grep("en", unique(df$myaxis), value = TRUE),
                           grep("vu", unique(df$myaxis), value = TRUE),
                           grep("nt", unique(df$myaxis), value = TRUE),
                           grep("lc", unique(df$myaxis), value = TRUE),
                           grep("dd", unique(df$myaxis), value = TRUE),
                           grep("not_assessed", unique(df$myaxis), value = TRUE),
                           grep("unknown", unique(df$myaxis), value = TRUE)))

df$global_IUCN<-factor(df$global_IUCN, levels=c("cr", "en", "vu", "nt", "lc", "dd", "not_assessed", "unknown"))



# plot
df %>%
  ggplot(aes(x=myaxis, y=indicatorPM , fill=global_IUCN)) +
  geom_violin(width=1.5, linewidth = 0)  +
  geom_jitter(size=.5, width = 0.1) +
  xlab("") + ylab("Proportion of mantained populations") +
  coord_flip() +
  scale_fill_manual(values= IUCNcolors, # iucn color codes
                    breaks=c(levels(df$global_IUCN))) +
  theme_light() +
  ggtitle("global Red List") +
  theme(panel.border = element_blank(), legend.position="none", text= element_text(size=15))

```

By regional IUCN (this may make no sense for your country if there is no IUCN redlist assessments at regional level:

                    
```{r indicatorPM, echo=F, message=F, warning=F}
                  ## Regional IUCN
                  
                  ## prepare data
                  # add sampling size
                  sample_size <- indicators_full %>%
                    filter(!is.na(indicatorPM)) %>% 
                    group_by(regional_redlist) %>% summarize(num=n())
                  
                  # new df 
                  df<- indicators_full %>% 
                    filter(!is.na(indicatorPM)) %>% 
                    # add sampling size 
                    left_join(sample_size) %>%
                    mutate(myaxis = paste0(regional_redlist, " (n= ", num, ")"))
                  
                  # change order of levels so that they are in the desired order
                  df$myaxis<-factor(df$myaxis, 
                                    #grep is used below to get the sample size, which may change depending on the data
                                    levels=c(grep("re", unique(df$myaxis), value = TRUE),
                                             grep("cr", unique(df$myaxis), value = TRUE),
                                             grep("en", unique(df$myaxis), value = TRUE),
                                             grep("vu", unique(df$myaxis), value = TRUE),
                                             grep("nt", unique(df$myaxis), value = TRUE),
                                             grep("lc", unique(df$myaxis), value = TRUE),
                                             grep("dd", unique(df$myaxis), value = TRUE),
                                             grep("not_assessed", unique(df$myaxis), value = TRUE),
                                             grep("unknown", unique(df$myaxis), value = TRUE)))
                  
                  df$regional_redlist<-factor(df$regional_redlist, levels=c("re","cr", "en", "vu", "nt", "lc", "dd", "not_assessed", "unknown"))
                  
                  # plot
                  df %>%
                    ggplot(aes(x=myaxis, y=indicatorPM , fill=regional_redlist)) +
                    geom_violin(width=1, linewidth = 0)  +
                    geom_jitter(size=.5, width = 0.1) +
                    xlab("") + ylab("Proportion of mantained populations") +
                    coord_flip() +
                    scale_fill_manual(values= IUCNcolors_regional, # iucn color codes
                                      breaks=c(levels(df$regional_redlist))) +
                    theme_light() +
                    ggtitle("regional Red List") +
                    theme(panel.border = element_blank(), legend.position="none", text= element_text(size=15))
                  
                  
                  
```

### Proportion of maintained populations by endemicity

```{r, echo=F, message=F, warning=F}
# get sample size by desired category
sample_size <- indicators_full  %>%
                    filter(!is.na(indicatorPM)) %>% 
                    group_by(national_endemic) %>% summarize(num=n())

# plot
indicators_full  %>% 
  # add sampling size 
  left_join(sample_size) %>%
  mutate(myaxis = paste0(national_endemic, " (n= ", num, ")")) %>%

  # plot
  ggplot(aes(x=myaxis, y=indicatorPM , fill=national_endemic)) +
      geom_violin(width=1, linewidth = 0)  +
      geom_jitter(size=.5, width = 0.1) +
      xlab("") + ylab("Proportion of popuations with Ne > 500") +
      coord_flip() +
      theme_light() +
      theme(panel.border = element_blank(), legend.position="right", text= element_text(size=20))
```

### Plot Scatter plot of indicatorPM and extant pops

```{r, echo=F, message=F, warning=F}
indicators_full %>%
  filter(!is.na(indicatorPM)) %>%
  # filter outliers with too many pops
  # filter(n_extant_populations<200) %>%
  
  # plot
    ggplot(aes(x=n_extant_populations, y=indicatorPM, color=defined_populations)) +
    geom_point() +
    theme_light() +
    theme(legend.position = "bottom") +
    ylab("Proportion of maintained populations") +
    xlab("Number of mantained populations") +
    theme(text = element_text(size = 13))
```